---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
## 1.Loading and preprocessing the data

```{r, echo=TRUE}
Data <- read.csv("activity.csv", sep =",")
```

## 2. What is mean total number of steps taken per day?

#### 2.1 Calculate the total number of steps taken per day
```{r, echo=FALSE}
library(ggplot2)
StepsByDay <- tapply(Data$steps, Data$date, sum, na.rm=TRUE)
head(StepsByDay)
```
#### 2.2 Make a histogram of the total number of steps taken each day
```{r, echo=TRUE}
qplot(StepsByDay, xlab="Total number of steps per day", ylab="frequency (binwidth=500)", binwidth=500)
```
#### 2.3 Calculate and report the mean and median of the total number of steps taken per day
``` {r, echo=TRUE}
MeanStepByDay<- mean(StepsByDay)
MedianStepByDay<-median(StepsByDay)
```
The mean is `r MeanStepByDay` and the median is `r MedianStepByDay`.
##3.What is the average daily activity pattern?

#### 3.1 Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, echo=TRUE}
StepsByInterval <- aggregate (steps~interval, Data, mean, na.rm=TRUE)
g <- ggplot (StepsByInterval, aes(interval, steps))
g+geom_line() + labs (x= "5 minutes interval", y="Average number of steps")
```
#### 3.2 Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo=TRUE}
StepsByInterval[which.max(StepsByInterval$steps),]
```

##4 Imputing missing values

#### 4.1 Calculate and report the total number of missing values in the dataset
```{r, echo=TRUE}
Miss <- sum(is.na(Data$steps))
```
The total number of missing values is `r Miss`.
#### 4.2 Devise a strategy for filling in all of the missing values in the dataset.
```{r, echo=TRUE}
library(Hmisc)
```
#### 4.3 Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo=TRUE}
head(Data)
DataImputed <- Data
DataImputed$steps <- impute(Data$steps, fun=mean)
head(DataImputed)
```
#### 4.4 Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r, echo=TRUE}
StepsByDayImputed <- tapply(DataImputed$steps, DataImputed$date, sum)
qplot(StepsByDayImputed, xlab="Total number of steps per day (Imputed", 
      ylab="frequency (binwidth=500) (Imputed)", binwidth=500)
```
Do these values differ from the estimates from the first part of the assignment? 
```{r, echo=TRUE}
MeanStepByDayImputed<- mean(StepsByDayImputed)
MedianStepByDayImputed <- median(StepsByDayImputed)
DiffMean <- MeanStepByDayImputed - MeanStepByDay
DiffMedian <- MedianStepByDayImputed - MedianStepByDay
DiffMean
DiffMedian
```
Yes, they differ. 
What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r, echo=TRUE}
DiffSteps <- sum(DataImputed$steps) - sum(Data$steps, na.rm=TRUE)
DiffSteps
```
The total number of steps is higer when the missing data are filled in.
