---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data



## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
## 1.Loading and preprocessing the data

```{r, echo=TRUE}
Data <- read.csv("activity.csv", sep =",")
```

## 2. What is mean total number of steps taken per day?

#### 2.1 Calculate the total number of steps taken per day
```{r, echo=FALSE}
library(ggplot2)
StepsByDay <- tapply(Data$steps, Data$date, sum, na.rm=TRUE)
head(StepsByDay)
```
#### 2.2 Make a histogram of the total number of steps taken each day
```{r, echo=TRUE}
qplot(StepsByDay, xlab="Total number of steps per day", ylab="frequency (binwidth=500)", binwidth=500)
```
#### 2.3 Calculate and report the mean and median of the total number of steps taken per day
``` {r, echo=TRUE}
MeanStepByDay<- mean(StepsByDay)
MedianStepByDay<-median(StepsByDay)
```
The mean is `r MeanStepByDay` and the median is `r MedianStepByDay`.
