---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## 1. Loading and preprocessing the data

```{r, echo=TRUE}
Data <- read.csv("activity.csv", sep =",")
```

## 2. What is mean total number of steps taken per day?

#### 2.1 Calculate the total number of steps taken per day
```{r, echo=FALSE}
library(ggplot2)
StepsByDay <- tapply(Data$steps, Data$date, sum, na.rm=TRUE)
head(StepsByDay)
```
#### 2.2 Make a histogram of the total number of steps taken each day
```{r, echo=TRUE}
qplot(StepsByDay, xlab="Total number of steps per day", ylab="frequency (binwidth=500)", binwidth=500)
```

#### 2.3 Calculate and report the mean and median of the total number of steps taken per day
``` {r, echo=TRUE}
MeanStepByDay<- mean(StepsByDay)
MedianStepByDay<-median(StepsByDay)
```
The mean is `r MeanStepByDay` and the median is `r MedianStepByDay`.
##3.What is the average daily activity pattern?

#### 3.1 Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, echo=TRUE}
StepsByInterval <- aggregate (steps~interval, Data, mean, na.rm=TRUE)
g <- ggplot (StepsByInterval, aes(interval, steps))
g+geom_line() + labs (x= "5 minutes interval", y="Average number of steps")
```

#### 3.2 Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo=TRUE}
StepsByInterval[which.max(StepsByInterval$steps),]
```

##4 Imputing missing values

#### 4.1 Calculate and report the total number of missing values in the dataset
```{r, echo=TRUE}
Miss <- sum(is.na(Data$steps))
```
The total number of missing values is `r Miss`.
#### 4.2 Devise a strategy for filling in all of the missing values in the dataset.
```{r, echo=TRUE}
library(Hmisc)
```
#### 4.3 Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo=TRUE}
head(Data)
DataImputed <- Data
DataImputed$steps <- impute(Data$steps, fun=mean)
head(DataImputed)
```
#### 4.4 Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r, echo=TRUE}
StepsByDayImputed <- tapply(DataImputed$steps, DataImputed$date, sum)
qplot(StepsByDayImputed, xlab="Total number of steps per day (Imputed", 
      ylab="frequency (binwidth=500) (Imputed)", binwidth=500)
```

Do these values differ from the estimates from the first part of the assignment? Yes, they differ. 
```{r, echo=TRUE}
MeanStepByDayImputed<- mean(StepsByDayImputed)
MedianStepByDayImputed <- median(StepsByDayImputed)
DiffMean <- MeanStepByDayImputed - MeanStepByDay
DiffMedian <- MedianStepByDayImputed - MedianStepByDay
DiffMean
DiffMedian
```
What is the impact of imputing missing data on the estimates of the total daily number of steps? The total number of steps is higer when the missing data are filled in.
```{r, echo=TRUE}
DiffSteps <- sum(DataImputed$steps) - sum(Data$steps, na.rm=TRUE)
DiffSteps
```
##5.Are there differences in activity patterns between weekdays and weekends?

#### 5.1 Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r, echo=TRUE}
DataImputed$Week <- ifelse(as.POSIXlt(DataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```
#### 5.2 Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r, echo=TRUE}
StepsByIntervalWeek <- aggregate(steps~interval+Week, DataImputed, mean)
g <- ggplot (StepsByIntervalWeek, aes(interval,steps))
g + geom_line()+ facet_grid(Week ~.) + labs (x<- "5 minutes iterval", 
                                             y="Average number of steps (imputed)")
```
